The Analogue Method for Precipitation Forecasting: Finding Better Analogue Situations at a Sub-Daily Time Step

### 2016-05-18 Moving temporal window - HESS/template.tex
Preparation for submission.

%% 2-Column Papers and Discussion Papers
%\documentclass[hess, manuscript]{copernicus}
\documentclass[hess]{copernicus}

@@

Previous tests showed the benefit, in terms of analogy criteria values, of searching for analogue synoptic situations at a finer time step, but without assessing the impact on the forecast skills \citep{Finet2008}. In order to assess this potential improvement of the prediction, precipitation data with a resolution of 10 minutes, on a respectable archive length, was used (section \ref{sec:data}). It was then aggregated in the form of 24~h-totals, but starting at different 6-hourly time step, by means of a moving total (section \ref{sec:method}).

@@

The changes due to the introduction of a STW was assessed first in regards to the analogy criteria improvement (section \ref{sec:influence_criteria}), and then in terms of precipitation prediction skill (section \ref{sec:influence_scores}). Finally, attempts were made to reconstruct longer 24~h-total precipitations series (section \ref{sec:attempts}).

@@

The study area is the upper Rh\^{o}ne catchment in Switzerland \cite[see also][]{Horton2012a}. Due to the low density of weather stations with high temporal resolution and long archives, no spatially aggregated rainfall was processed. The time series (on the period 1982-2007) come from 6 automatic weather stations, namely Ulrichen, Zermatt, Visp, Montana, Sion and Aigle (Figure \ref{fig:map}), which are subject to various meteorological influences \citep{Horton2012}. The results will thereafter be presented arbitrarily for the Ulrichen station, but they are equivalent for all stations. The precipitation time series were then aggregated by means of a moving 24~h-total (see next section).

@@

In order to assess the benefit of searching analogue situations at a sub-daily time step, an appropriate precipitation series is required. On the basis of high resolution time series (section \ref{sec:data}), 24~h-totals were processed, but at a 6-hourly time step (temporal resolution of the reanalysis dataset), by means of a moving 24~h-total.

The target situations and their corresponding observed precipitation values (used for validation) do not change, because the forecast is always established for a fixed period of the target day (6-30~h), as before. The difference is that the candidates are 4~times more numerous (Figure \ref{fig:principle}).

@@

The improvements of the S1 criteria are summarized in Figure \ref{fig:changes_S1}, which shows (top) quantiles of the S1 criteria according to the analogue rank for the conventional method and the STW, and (bottom) quantiles of the relative reduction. This confirms that all quantile seem similarly reduced (S1 distributions keep their shape), and that this improvement is constantly increasing from the first to the last analogue (Figure \ref{fig:changes_S1} bottom).

The median of the S1 values reduction (Figure \ref{fig:changes_S1} bottom) starts approximately at 5~\% for the first analogue and reaches more than 10~\% for the last one. This increasing trend with the analogue rank can be explained by the accumulation of better analogues in the distribution. The minimum improvement starts from 0 and reaches 2-3\%, meaning that the criteria have been improved on most analogues for every day of the calibration period. All other stations have a similar improvement of the S1 criteria, both in terms of distribution shape and amplitude.

@@

The dynamism of a given atmospheric condition cannot be easily quantified. A basic assumption is considered here, which stipulates that the more a day is rainy, the more dynamic the situation is. The results of this analysis are summarized in Figure \ref{fig:changes_S1_precip_threshold} by the median reduction of S1 for days with precipitation (organized into classes) between two thresholds. The number of cases per class being reduced, the curves are not as smooth as in previous analyzes. It is nevertheless clear that the improvement tends to increase on days with higher precipitation. This is true for all our stations and confirms our intuition.

@@

A seasonal effect can be observed on the reduction of the S1 criteria due to the STW (Figure \ref{fig:changes_S1_seasons}). The improvements are greater for winter than summer. One hypothesis is that the diurnal effects of the summer months have an influence on the atmospheric circulation, at least in the lower layers. This effect is based on the daily cycle and good analogues are essentially found for the same hours or time windows.

An analysis of the selected hours for the geopotential predictor seems to confirm this assumption (Figure \ref{fig:hours_selection_per_season}). It was found that the new choice of the temporal window in winter, when using the STW approach, is well balanced between the 4 options. This means a change of 75\% of the analogues selection compared to the conventional approach, which improves the circulation analogy.

On the contrary, the summer months have a preference for the initial temporal window (Z500 24h \& Z1000 12h), due to more pronounced diurnal effects which reduces the potential for improvement of the criteria. Other seasons are between these two extremes, which is consistent with their respective improvements. This seasonal effect was observed for each station in a very similar way, and generally even with a slightly larger amplitude than for Ulrichen (see example of the Visp station in Figure \ref{fig:hours_selection_per_season_Visp}).

@@

When considering the second level of analogy of the 2Z-2MI method (Table \ref{table:method_2Z-2MI}), the number of candidates situations did not increase, as they are subsampled in the previously $N_{1}$ selected analogues, but their dates have changed. In contrast to earlier, both a reduction or an increase of the RMSE analogy criterion values are possible. Indeed, Figure \ref{fig:changes_RMSE} shows an almost insignificant improvement of the RMSE values. Unlike for the first level of analogy, the relative improvement of the RMSE values are distributed relatively symmetrically around zero, with improvements and losses of the same amplitude. Once again, the results for the other stations are similar.

@@

A systematic improvement of the S1 values was previously observed. However, finding better analogue situations does not obligatory imply better skills to predict the precipitation. Therefore, the impact of the STW introduction, and thus the selection of other analogue dates, has to be assessed on the performance scores. In order to perform this assessment, the 24~h-totals (moving average) at a 6-hourly time step were used (see section \ref{sec:data}). The target dates remain unchanged, since the original time slot (6~h~UTC - 6~h~UTC the next day) is kept, and thus the performance scores can be directly compared with the former ones.

@@

The S1 criteria was previously found to be improved to a greater extent for the most dynamic situations related with higher precipitation values (section \ref{sec:influence_precip}). The changes in terms of performance scores will now be assessed regarding precipitation thresholds. Figure \ref{fig:changes_CRPS_precip_threshold} synthesizes these differences for the Ulrichen station, other stations having the same behavior.

@@

After recalibrating, some changes in the optimal parameters can be observed for both 2Z (Table \ref{table:params_2Z_new}) and 2Z-2MI methods (Table \ref{table:params_2Z-2MI_new}). Among these, the east-west dimension of the spatial windows of the circulation analogy tends to decrease. More significantly, the optimal number of analogues increases after introducing the STW, of a significant order of magnitude: 25~\% to 83~\% for the 2Z method and 20~\% 67~\% for the 2Z-2MI method. The number of analogues of the first analogy level of the 2Z-2MI method even reached three times its previous value for the Visp station. It seems as if the method perceive the increase in the archive length and diversity by allowing to extract more analogues. As we saw in Figure \ref{fig:changes_S1}, the improvement of the S1 criteria grows along with the rank of the analogue, which shows an accumulation of better analogue situations in the distributions. It seems thus profitable to widen the selection of analogues in order to keep also some whose rank has increased, as they appear to be also relevant to predict precipitation values. The number of good analogues is thus globally increased.

This increase in number of analogues has a slight effect on the performance of the different precipitation classes. The same analysis as in section \ref{sec:improvement_CRPSS_precip_threshold} has been performed again on the basis of the newly calibrated parameters. The results are generally very similar, but a slight performance increase of small precipitation values can be observed at the expense of higher amounts. The change in the analogues numbers is likely to be responsible for this difference in the performance distributions.

@@

It appears that the sharpness is a bit inferior with the STW in favor of accuracy (Table \ref{table:CRPSS_decomposition_no_recalibration}). When using the recalibrated parameters, the components are a bit more balanced, but still the accuracy prevails on the sharpness (Table \ref{table:CRPSS_decomposition_recalibration}). Figure \ref{fig:CRPSS_decomposition_precip_thresholds} illustrates the changes (relative to the total CRPS) in sharpness and accuracy for different precipitation thresholds at the Ulrichen station. Be it for the 2Z or 2Z-2MI method, the original or the recalibrated parameters, the changes are similar: the improvements concern the accuracy over the sharpness, and this to a greater extent for significant precipitation events. The STW does not allow to improve the sharpness, but the improvement in accuracy is significant, and this especially for the days with heavy precipitation. In terms of the predicted precipitation distribution, it means that the median of the prediction is closer to the observed values, while the distribution is not more condensed than previously.

@@

A proportional distribution is certainly the simplest approach that can be performed. It consists in allocating proportional parts of the original daily time series into a new moving average of 24~h-totals (Figure \ref{fig:illustration_disaggregation}). When using this reconstructed time series on the shorter period, the method performance was degraded (Table \ref{table:disaggregation_proportional} to compare to Table \ref{table:CRPSS_STW}) and was even below the conventional method without STW (Table \ref{table:loss_reduction}). The benefit of a better selection of the analogue situations is lost due to a precipitation archive of poor quality.

\begin{table}[htb]
	\caption{Values of the CRPSS (\%) skill score for the original and the recalibrated parameters (with the sequential method, as described in Section \ref{sec:introduction}) using the STW approach on the disaggregated precipitation time series (short period) by means of the proportional distribution.}
	\begin{center}
		\begin{tabular}{l c c c c}
			\hline
			\multirow{2}{*}{Station} & \multicolumn{2}{c}{2Z} & \multicolumn{ 2}{c}{2Z-2MI} \\
			& original & recalib. & original & recalib. \\
			\hline
			Ulrichen & 29.13 & 29.61 & 33.15 & 33.45 \\
			Zermatt & 22.17 & 22.80 & 26.72 & 27.43 \\
			Visp & 22.32 & 22.89 & 27.01 & 28.04 \\
			Montana & 29.41 & 30.24 & 33.83 & 34.55 \\
			Sion & 22.98 & 23.41 & 28.57 & 29.15 \\
			Aigle & 29.07 & 29.46 & 34.66 & 35.09 \\
			\hline
		\end{tabular}
	\end{center}
	\label{table:disaggregation_proportional}
\end{table}

@@

As demonstrated in the previous section, the performance improvement brought by the STW is lost when using a simplistic precipitation time series. In order to get closer to the chronology of the actual precipitation, the reconstruction procedure must rely on some informative variable. Data from the NCEP/NCAR reanalysis 1 (section \ref{sec:data}) was used in this perspective, despite their rough resolution.

@@

\begin{table}[htb]
	\caption{Values of the CRPSS (\%) skill score for Zermatt, for the original and the recalibrated parameters (with the sequential method, as described in Section \ref{sec:introduction}) using the STW approach on the disaggregated precipitation time series (short and long periods) by means of variables from the reanalysis.}
	\begin{center}
		\begin{tabular}{l c c c c}
			\hline
			\multirow{2}{*}{Period} & \multicolumn{2}{c}{2Z} & \multicolumn{ 2}{c}{2Z-2MI} \\
			& original & recalib. & original & recalib. \\
			\hline
			1982-2007 & 22.57 & 23.14 & 27.11 & 27.71 \\
			1961-2008 & 23.81 & 24.38 & 28.42 & 28.86 \\
			\hline
		\end{tabular}
	\end{center}
	\label{table:proxy_CRPSS}
\end{table}

@@

>>>>

%% 2-Column Papers and Discussion Papers
\documentclass[hess, manuscript]{copernicus}
%\documentclass[hess]{copernicus}

@@

Previous tests showed the benefit, in terms of analogy criteria values, of searching for analogue synoptic situations at a finer time step, but without assessing the impact on the forecast skills \citep{Finet2008}. In order to assess this potential improvement of the prediction, precipitation data with a resolution of 10 minutes, on a respectable archive length, was used (Sect. \ref{sec:data}). It was then aggregated in the form of 24~h-totals, but starting at different 6-hourly time step, by means of a moving total (Sect. \ref{sec:method}).

@@

The changes due to the introduction of a STW was assessed first in regards to the analogy criteria improvement (Sect. \ref{sec:influence_criteria}), and then in terms of precipitation prediction skill (Sect. \ref{sec:influence_scores}). Finally, attempts were made to reconstruct longer 24~h-total precipitations series (Sect. \ref{sec:attempts}).

@@

The study area is the upper Rh\^{o}ne catchment in Switzerland \cite[see also][]{Horton2012a}. Due to the low density of weather stations with high temporal resolution and long archives, no spatially aggregated rainfall was processed. The time series (on the period 1982-2007) come from 6 automatic weather stations, namely Ulrichen, Zermatt, Visp, Montana, Sion and Aigle (Fig. \ref{fig:map}), which are subject to various meteorological influences \citep{Horton2012}. The results will thereafter be presented arbitrarily for the Ulrichen station, but they are equivalent for all stations. The precipitation time series were then aggregated by means of a moving 24~h-total (see next section).

@@

In order to assess the benefit of searching analogue situations at a sub-daily time step, an appropriate precipitation series is required. On the basis of high resolution time series (Sect. \ref{sec:data}), 24~h-totals were processed, but at a 6-hourly time step (temporal resolution of the reanalysis dataset), by means of a moving 24~h-total.

The target situations and their corresponding observed precipitation values (used for validation) do not change, because the forecast is always established for a fixed period of the target day (6-30~h), as before. The difference is that the candidates are 4~times more numerous (Fig. \ref{fig:principle}).

@@

The improvements of the S1 criteria are summarized in Fig. \ref{fig:changes_S1}, which shows (top) quantiles of the S1 criteria according to the analogue rank for the conventional method and the STW, and (bottom) quantiles of the relative reduction. This confirms that all quantile seem similarly reduced (S1 distributions keep their shape), and that this improvement is constantly increasing from the first to the last analogue (Fig. \ref{fig:changes_S1} bottom).

The median of the S1 values reduction (Fig. \ref{fig:changes_S1} bottom) starts approximately at 5~\% for the first analogue and reaches more than 10~\% for the last one. This increasing trend with the analogue rank can be explained by the accumulation of better analogues in the distribution. The minimum improvement starts from 0 and reaches 2-3\%, meaning that the criteria have been improved on most analogues for every day of the calibration period. All other stations have a similar improvement of the S1 criteria, both in terms of distribution shape and amplitude.

@@

The dynamism of a given atmospheric condition cannot be easily quantified. A basic assumption is considered here, which stipulates that the more a day is rainy, the more dynamic the situation is. The results of this analysis are summarized in Fig. \ref{fig:changes_S1_precip_threshold} by the median reduction of S1 for days with precipitation (organized into classes) between two thresholds. The number of cases per class being reduced, the curves are not as smooth as in previous analyzes. It is nevertheless clear that the improvement tends to increase on days with higher precipitation. This is true for all our stations and confirms our intuition.

@@

A seasonal effect can be observed on the reduction of the S1 criteria due to the STW (Fig. \ref{fig:changes_S1_seasons}). The improvements are greater for winter than summer. One hypothesis is that the diurnal effects of the summer months have an influence on the atmospheric circulation, at least in the lower layers. This effect is based on the daily cycle and good analogues are essentially found for the same hours or time windows.

An analysis of the selected hours for the geopotential predictor seems to confirm this assumption (Fig. \ref{fig:hours_selection_per_season}). It was found that the new choice of the temporal window in winter, when using the STW approach, is well balanced between the 4 options. This means a change of 75\% of the analogues selection compared to the conventional approach, which improves the circulation analogy.

On the contrary, the summer months have a preference for the initial temporal window (Z500 24h \& Z1000 12h), due to more pronounced diurnal effects which reduces the potential for improvement of the criteria. Other seasons are between these two extremes, which is consistent with their respective improvements. This seasonal effect was observed for each station in a very similar way, and generally even with a slightly larger amplitude than for Ulrichen (see example of the Visp station in Fig. \ref{fig:hours_selection_per_season_Visp}).

@@

When considering the second level of analogy of the 2Z-2MI method (Table \ref{table:method_2Z-2MI}), the number of candidates situations did not increase, as they are subsampled in the previously $N_{1}$ selected analogues, but their dates have changed. In contrast to earlier, both a reduction or an increase of the RMSE analogy criterion values are possible. Indeed, Fig. \ref{fig:changes_RMSE} shows an almost insignificant improvement of the RMSE values. Unlike for the first level of analogy, the relative improvement of the RMSE values are distributed relatively symmetrically around zero, with improvements and losses of the same amplitude. Once again, the results for the other stations are similar.

@@

A systematic improvement of the S1 values was previously observed. However, finding better analogue situations does not obligatory imply better skills to predict the precipitation. Therefore, the impact of the STW introduction, and thus the selection of other analogue dates, has to be assessed on the performance scores. In order to perform this assessment, the 24~h-totals (moving average) at a 6-hourly time step were used (see Sect. \ref{sec:data}). The target dates remain unchanged, since the original time slot (6~h~UTC - 6~h~UTC the next day) is kept, and thus the performance scores can be directly compared with the former ones.

@@

The S1 criteria was previously found to be improved to a greater extent for the most dynamic situations related with higher precipitation values (Sect. \ref{sec:influence_precip}). The changes in terms of performance scores will now be assessed regarding precipitation thresholds. Fig. \ref{fig:changes_CRPS_precip_threshold} synthesizes these differences for the Ulrichen station, other stations having the same behavior.

@@

After recalibrating, some changes in the optimal parameters can be observed for both 2Z (Table \ref{table:params_2Z_new}) and 2Z-2MI methods (Table \ref{table:params_2Z-2MI_new}). Among these, the east-west dimension of the spatial windows of the circulation analogy tends to decrease. More significantly, the optimal number of analogues increases after introducing the STW, of a significant order of magnitude: 25~\% to 83~\% for the 2Z method and 20~\% 67~\% for the 2Z-2MI method. The number of analogues of the first analogy level of the 2Z-2MI method even reached three times its previous value for the Visp station. It seems as if the method perceive the increase in the archive length and diversity by allowing to extract more analogues. As we saw in Fig. \ref{fig:changes_S1}, the improvement of the S1 criteria grows along with the rank of the analogue, which shows an accumulation of better analogue situations in the distributions. It seems thus profitable to widen the selection of analogues in order to keep also some whose rank has increased, as they appear to be also relevant to predict precipitation values. The number of good analogues is thus globally increased.

This increase in number of analogues has a slight effect on the performance of the different precipitation classes. The same analysis as in Sect. \ref{sec:improvement_CRPSS_precip_threshold} has been performed again on the basis of the newly calibrated parameters. The results are generally very similar, but a slight performance increase of small precipitation values can be observed at the expense of higher amounts. The change in the analogues numbers is likely to be responsible for this difference in the performance distributions.

@@

It appears that the sharpness is a bit inferior with the STW in favor of accuracy (Table \ref{table:CRPSS_decomposition_no_recalibration}). When using the recalibrated parameters, the components are a bit more balanced, but still the accuracy prevails on the sharpness (Table \ref{table:CRPSS_decomposition_recalibration}). Fig. \ref{fig:CRPSS_decomposition_precip_thresholds} illustrates the changes (relative to the total CRPS) in sharpness and accuracy for different precipitation thresholds at the Ulrichen station. Be it for the 2Z or 2Z-2MI method, the original or the recalibrated parameters, the changes are similar: the improvements concern the accuracy over the sharpness, and this to a greater extent for significant precipitation events. The STW does not allow to improve the sharpness, but the improvement in accuracy is significant, and this especially for the days with heavy precipitation. In terms of the predicted precipitation distribution, it means that the median of the prediction is closer to the observed values, while the distribution is not more condensed than previously.

@@

A proportional distribution is certainly the simplest approach that can be performed. It consists in allocating proportional parts of the original daily time series into a new moving average of 24~h-totals (Fig. \ref{fig:illustration_disaggregation}). When using this reconstructed time series on the shorter period, the method performance was degraded (Table \ref{table:disaggregation_proportional} to compare to Table \ref{table:CRPSS_STW}) and was even below the conventional method without STW (Table \ref{table:loss_reduction}). The benefit of a better selection of the analogue situations is lost due to a precipitation archive of poor quality.

@@

As demonstrated in the previous section, the performance improvement brought by the STW is lost when using a simplistic precipitation time series. In order to get closer to the chronology of the actual precipitation, the reconstruction procedure must rely on some informative variable. Data from the NCEP/NCAR reanalysis 1 (Sect. \ref{sec:data}) was used in this perspective, despite their rough resolution.

@@

@@

\clearpage

\begin{table}[htb]
	\caption{Values of the CRPSS (\%) skill score for the original and the recalibrated parameters (with the sequential method, as described in Sect. \ref{sec:introduction}) using the STW approach on the disaggregated precipitation time series (short period) by means of the proportional distribution.}
	\begin{center}
		\begin{tabular}{l c c c c}
			\hline
			\multirow{2}{*}{Station} & \multicolumn{2}{c}{2Z} & \multicolumn{ 2}{c}{2Z-2MI} \\
			& original & recalib. & original & recalib. \\
			\hline
			Ulrichen & 29.13 & 29.61 & 33.15 & 33.45 \\
			Zermatt & 22.17 & 22.80 & 26.72 & 27.43 \\
			Visp & 22.32 & 22.89 & 27.01 & 28.04 \\
			Montana & 29.41 & 30.24 & 33.83 & 34.55 \\
			Sion & 22.98 & 23.41 & 28.57 & 29.15 \\
			Aigle & 29.07 & 29.46 & 34.66 & 35.09 \\
			\hline
		\end{tabular}
	\end{center}
	\label{table:disaggregation_proportional}
\end{table}

\begin{table}[htb]
	\caption{Values of the CRPSS (\%) skill score for Zermatt, for the original and the recalibrated parameters (with the sequential method, as described in Sect. \ref{sec:introduction}) using the STW approach on the disaggregated precipitation time series (short and long periods) by means of variables from the reanalysis.}
	\begin{center}
		\begin{tabular}{l c c c c}
			\hline
			\multirow{2}{*}{Period} & \multicolumn{2}{c}{2Z} & \multicolumn{ 2}{c}{2Z-2MI} \\
			& original & recalib. & original & recalib. \\
			\hline
			1982-2007 & 22.57 & 23.14 & 27.11 & 27.71 \\
			1961-2008 & 23.81 & 24.38 & 28.42 & 28.86 \\
			\hline
		\end{tabular}
	\end{center}
	\label{table:proxy_CRPSS}
\end{table}
